---
title: "Homework 2"
author: "Marisa Sobel"
date: "9/28/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1

```{r, message=FALSE, collapse=TRUE}
# import transit data, clean, select relevant vars, recode 'entry'

nyctransit_data = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, 
         station_longitude, starts_with("route"), 
         entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))

nyctransit_data
```

### About `nyctransit_data`

The tibble created has **`r nrow(nyctransit_data)`** observations and **`r ncol(nyctransit_data)`** columns. It includes variables that describe the following: line (`line`), station name (`station_name`), station latitude (`station_latitude`), station longitude (`station_longitude`), entrance type (`entrance_type`), entry (`entry`), vending (`vending`), ADA compliance (`ada`), and all of the different route numbers. Most are variables with individual values except for the `route(#)` columns. This means that the data are not tidy, as there are missing values marked as "NA" in the `route(#)` columns. 

### Questions

1. _*How many distict stations are there?*_ 

```{r}
# distinct statations by 'line' and 'station name'
distinct(nyctransit_data, line, station_name)
```

There are **465 distinct stations**, identified by both station name and line. 

2. _*How many stations are ADA compliant?*_ 

```{r}
# distinct stations with ADA compliance
distinct(nyctransit_data, line, station_name, .keep_all = TRUE) %>% 
  filter(ada == TRUE)
```

**84 stations** are ADA compliant. 

3. _*What proportion of station entrances/exits without vending allow entrence?*_ 

```{r}
# proportion entrances/exits w/o vending that allow entry
filter(nyctransit_data, vending == "NO") %>% 
  filter(entry == TRUE)
```

**69/183 stations** or **37% of stations** without vending allow entrance. 

### Reformat

_*How many distinct stations serve the A train? Of stations that serve the A train, how many are ADA compliant?*_ 

```{r, collapse=TRUE}
# wide to long format for 'route', new key = 'route_number'
# filter for A train, then distinct stations
# of those, number many ADA compliant

nyctransit_data_tidy = 
  gather(nyctransit_data, key = route_number,
         value = route_name, route1:route11) %>% 
  filter(route_name == "A") %>% 
  distinct(line, station_name, .keep_all = TRUE) %>% 
  filter(ada == TRUE)

nyctransit_data_tidy
```

**60 distinct stations** serve the A train. **17** of those stations are ADA compliant. 


## Problem 2 

Mr. Trash Wheel Data

```{r, collapse=TRUE}
# import trash wheel data with relevant columbs (no notes) clean names
# remove rows without dumpster data, remove final sum at end
# make sports_balls numeric 

trashwheel_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
             sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster), dumpster != 23220) %>% 
  mutate(sports_balls = as.integer(sports_balls))

trashwheel_data
```

Separate precipitation data

```{r, collapse=TRUE}
# import 2016 precip data, clean names, rename col, remove NAs, make all col numeric, add 'year' var

precip_2016 = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
             sheet = "2016 Precipitation") %>% 
  janitor::clean_names() %>% 
  rename(month = precipitation_in, total_precip = x_1) %>%
  filter(!is.na(month), !is.na(total_precip),
         month != "Month", total_precip != "Total") %>% 
  mutate(
    month = as.numeric(month),
    total_precip = as.numeric(total_precip),
    year = "2016")

head(precip_2016)

# import 2017 precip data, repeate above

precip_2017 = 
  read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", 
             sheet = "2017 Precipitation") %>% 
  janitor::clean_names() %>% 
  rename(month = precipitation_in, total_precip = x_1) %>% 
  filter(!is.na(month), !is.na(total_precip), 
         month != "Month", total_precip != "Total") %>% 
  mutate(
    month = as.numeric(month),
    total_precip = as.numeric(total_precip),
    year = "2017")

precip_2017
```

Combine precipitation data

```{r, collapse=TRUE}
# bind cleaned 2016 and 2017 precip tibbles, numeric months -> names
precip_tidy = bind_rows(precip_2016, precip_2017) %>% 
  mutate(month = month.name[month]) 

precip_tidy
```

### About `trashwheel_data` and `precip_tidy`

```{r, collapse=TRUE}
# median sports balls in 2016
trashwheel_data %>% 
  group_by(year) %>% 
  filter(year == "2016") %>% 
  summarise(median(sports_balls))

# total precip in 2016
precip_tidy %>% 
  group_by(year) %>% 
  filter(year == "2016") %>% 
  summarise(sum(total_precip))

# total precip in 2017
precip_tidy %>% 
  group_by(year) %>% 
  filter(year == "2017") %>% 
  summarise(sum(total_precip))
```

There are **`r nrow(trashwheel_data)`** observations in the Mr. Trash Wheel dataset, and **`r nrow(precip_tidy)`** observations  in the 2016-2017 monthly percipitation dataset. From `r trashwheel_data$month[1]` `r trashwheel_data$year[1]` until `r trashwheel_data$month[215]` `r trashwheel_data$year[215]`, Mr. Trash Wheel collected **`r sum(trashwheel_data$weight_tons)`** tons or **`r sum(trashwheel_data$volume_cubic_yards)`** cubic yards of trash from the Baltimore Harbor. Over the same course of time, the dumpster monthly averages of some of the objects counted were as followed: **`r round(mean(trashwheel_data$plastic_bottles))` plastic bottles**, **`r round(mean(trashwheel_data$polystyrene))` polystyrene**, and **`r round(mean(trashwheel_data$cigarette_butts))` cigarette butts**. In 2017, the median number of sports balls in a dumpster was **26 balls**. For the available data, the total precipitation was **29.9 inches** in 2017 and **40.0 inches** in 2016. 


## Problem 3 

```{r, collapse=TRUE}
# add library that includes dataset 
library("p8105.datasets")

# import brfss_smart2010 data, clean names, filter by "Overall Health", remove unwanted vars
# long -> wide with key=reponse and value= data_value
# new var: proportion (excellent + very good / total)

brfss_data = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  select(-class, -topic, -question, 
         -sample_size, -(confidence_limit_low:geo_location)) %>% 
  spread(key = response, value = data_value) %>% 
  janitor::clean_names() %>% 
  mutate(
    prop_high_positive = (excellent + very_good) /
      (excellent + very_good + good + fair + poor))

brfss_data
```


